Natural Language Interfaces to Structured Data

### 2012-08-06 tr-nlisurvey/src/nli-survey.tex
commit

Question typing is a research task within the Q\&A field.
This community has classified questions in \emph{facto\"id questions} (i.e.
questions that can be answered by an \emph{objective} fact, typically questions
staring with wh-words except ``why'') and \emph{complex questions} which
answers are usually \emph{subjective} to the answerer, and for which there
might be several \emph{true} answers, possibly contradicting each other.
Complex questions can be ``why''-questions, ``how''-questions or
\emph{definition}-questions.
Facto\"id question is defined by Soricut \&
Brill~\cite{Soricut:2006:AQA:1127331.1127342} as questions ``for which a
complete answer can be given in 50 bytes or less, which is roughly a few
words''.
Several finer classification have been proposed in specific domains or
applications. 
For instance, in the case of temporal questions, Saquete et
al.~\cite{Saquete:2004:SCT:1218955.1219027} have proposed the classification
reproduced table~\ref{tab:saquete-classification}.
\begin{table}
\centering
\begin{tabular}{ll}\hline
\multicolumn{1}{c}{\textbf{Question type}} &
\multicolumn{1}{c}{\textbf{Example}}\\\hline\hline
Single event temporal questions &
``When did Jordan close \\
 without temporal expressions & the port of Aqaba to Kuwait?''\\\hline 
 Single event temporal questions & 
``Who won the 1988 New \\
 with temporal expressions & Hampshire republican primary?''\\\hline
 \multirow{2}{*}{Multiple events temporal questions} & ``What did G. Bush do
 after\\
  & the U.N. Security Council\\
 \multirow{2}{*}{with temporal expressions} & ordered a global embargo on\\
  & trade with Irak in August 90?''\\\hline
  \multirow{2}{*}{Multiple events temporal questions} & ``What happened to world
  oil\\
   & prices after the Iraqi\\
  without temporal expressions & annexation of Kuwait?''\\\hline

@@

\subsection{Database queries}
\label{sec:big-picture-semantic}
\TODO{Discuss query languages; discuss the different types of queries expressed
by these languages, e.g. entity queries, relational queries, additional
constructs such as group by etc.}

The parse tree (which may or not contain semantic information in parse tree
nodes) must then be {\it interpreted} in some internal semantic representation.
Table~\ref{tab:semantic-meaning-representation} displays semantic meaning
representations used in different systems.
\begin{table*}[!h]
\centering
\begin{tabular}{|c|c|c|}\hline
{\bf System} & {\bf Internal representation} & {\bf DB query}\\\hline\hline
{\sc Baseball}~\cite{Green:1961:BAQ:1460690.1460714} & specification list &
$\surd$\\\hline
{\sc Lunar}~\cite{Woods:1973:PNL:1499586.1499695} & meaning representation
language & X\\\hline 
{\sc Chat-80}~\cite{Warren:1982:EEA:972942.972944} &
logic expression & X\\\hline
{\sc Team}~\cite{Grosz:1987:TED:25672.25674} & logic expression & X\\\hline
{\sc Qwerty}~\cite{Nelken:2000:QTD:992730.992808} & logic expression & X\\\hline
{\sc Irus}~\cite{Bates:1983:IRU:511793.511804} & meaning representation language
& X\\\hline
{\sc Precise}~\cite{Popescu:2003:TTN:604045.604070} & graph representation &
X\\\hline
{\sc Masque/SQL}~\cite{Androutsopoulos93masque} & meaning representation
language (LQL) & X \\\hline 
{\sc NaLIX} \& {\sc
DaNaLIX}~\cite{Li:2005:NIN:1066157.1066281,Li:2007:DDN:1247480.1247643} & XQuery
& $\surd$\\\hline 
{\sc C-Phrase}~\cite{Minock:2010:CSB:1715942.1716190} & $\lambda$-calculus & X
\\\hline
{\sc Panto}~\cite{Wang:2007:PPN:1419662.1419706} & graph representation (query
triples) & X\\\hline 
{\sc Orakel}~\cite{Cimiano:2007:PNL:1216295.1216330} & $\lambda$-calculus &
X\\\hline
Miller et al.~\cite{Miller:1996:FSA:981863.981871} &
semantic frames & X\\\hline 
Zettlemoyer et
Collins~\cite{DBLP:conf/uai/ZettlemoyerC05} & $\lambda$-calculus &
$\surd$\\\hline 
{\sc Wolfie}~\cite{Thompson:2003:AWM:1622420.1622421} & Prolog & $\surd$
\\\hline {\sc PowerAqua}~\cite{DBLP:conf/esws/LopezMU06} & graph representation
(query triples) & X\\\hline
{\sc DeepQA}~\cite{FerrucciBCFGKLMNPSW10} & semantic triples & $\surd$\\\hline
\end{tabular}
\caption{Semantic meaning representations}
\label{tab:semantic-meaning-representation}
\end{table*}
As shown in this table, the internal semantic representation can be or not the
targetted query representation. 
The expressivity of the database query is discussed in section
section~\ref{sec:big-picture-expressivity}. 
The semantic representation is intented to capture as much as possible users'
intent, and is sufficient to generate the final database query. 
While the syntactic representation may contain lots of ambiguities, the semantic
representation does not. 
The semantic representation depends on the data structure. However, to
demonstrate their portability features, some systems contain components that
can translate the semantic representation for different database query
languages.

@@

\subsection{Domain-independence}

@@

\subsubsection{Classic translation approach}
TOBECONSIDERED
----------
The classic translation approach consists in going from (1) to (4).
To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.

>>>>

Question typing is a research task within the Q\&A field.
This community has classified questions in \emph{facto\"id questions} (i.e.
questions that can be answered by an \emph{objective} fact, typically questions
staring with wh-words except ``why'') and \emph{complex questions} which
answers are usually \emph{subjective} to the answerer, and for which there
might be several \emph{true} answers, possibly contradicting each other.
Complex questions can be ``why''-questions, ``how''-questions or
\emph{definition}-questions.
Facto\"id question is defined by Soricut \&
Brill~\cite{Soricut:2006:AQA:1127331.1127342} as questions ``for which a
complete answer can be given in 50 bytes or less, which is roughly a few
words'' (see also Kwok et al. ``Scaling question answering to the Web, WWW).
Several finer classification have been proposed in specific domains or
applications. 
For instance, in the case of temporal questions, Saquete et
al.~\cite{Saquete:2004:SCT:1218955.1219027} have proposed the classification
reproduced table~\ref{tab:saquete-classification}.
\begin{table}
\centering
\begin{tabular}{ll}\hline
\multicolumn{1}{c}{\textbf{Question type}} &
\multicolumn{1}{c}{\textbf{Example}}\\\hline\hline
Single event temporal questions &
``When did Jordan close \\
 without temporal expressions & the port of Aqaba to Kuwait?''\\\hline 
 Single event temporal questions & 
``Who won the 1988 New \\
 with temporal expressions & Hampshire republican primary?''\\\hline
 \multirow{2}{*}{Multiple events temporal questions} & ``What did G. Bush do
 after\\
  & the U.N. Security Council\\
 \multirow{2}{*}{with temporal expressions} & ordered a global embargo on\\
  & trade with Irak in August 90?''\\\hline
  \multirow{2}{*}{Multiple events temporal questions} & ``What happened to world
  oil\\
   & prices after the Iraqi\\
  without temporal expressions & annexation of Kuwait?''\\\hline

@@

\subsection{Database queries}
\label{sec:big-picture-semantic}
%\TODO{Discuss query languages; discuss the different types of queries expressed
%by these languages, e.g. entity queries, relational queries, additional
%constructs such as group by etc.}

\begin{table*}[h!]
\centering
\begin{tabular}{|c|c|c|}\hline
{\bf System} & {\bf Internal representation} & {\bf DB query}\\\hline\hline
{\sc Baseball}~\cite{Green:1961:BAQ:1460690.1460714} & specification list &
$\surd$\\\hline
{\sc Lunar}~\cite{Woods:1973:PNL:1499586.1499695} & meaning representation
language & X\\\hline 
{\sc Chat-80}~\cite{Warren:1982:EEA:972942.972944} &
logic expression & X\\\hline
{\sc Team}~\cite{Grosz:1987:TED:25672.25674} & logic expression & X\\\hline
{\sc Qwerty}~\cite{Nelken:2000:QTD:992730.992808} & logic expression & X\\\hline
{\sc Irus}~\cite{Bates:1983:IRU:511793.511804} & meaning representation language
& X\\\hline
{\sc Precise}~\cite{Popescu:2003:TTN:604045.604070} & graph representation &
X\\\hline
{\sc Masque/SQL}~\cite{Androutsopoulos93masque} & meaning representation
language (LQL) & X \\\hline 
{\sc NaLIX} \& {\sc
DaNaLIX}~\cite{Li:2005:NIN:1066157.1066281,Li:2007:DDN:1247480.1247643} & XQuery
& $\surd$\\\hline 
{\sc C-Phrase}~\cite{Minock:2010:CSB:1715942.1716190} & $\lambda$-calculus & X
\\\hline
{\sc Panto}~\cite{Wang:2007:PPN:1419662.1419706} & graph representation (query
triples) & X\\\hline 
{\sc Orakel}~\cite{Cimiano:2007:PNL:1216295.1216330} & $\lambda$-calculus &
X\\\hline
Miller et al.~\cite{Miller:1996:FSA:981863.981871} &
semantic frames & X\\\hline 
Zettlemoyer et
Collins~\cite{DBLP:conf/uai/ZettlemoyerC05} & $\lambda$-calculus &
$\surd$\\\hline 
{\sc Wolfie}~\cite{Thompson:2003:AWM:1622420.1622421} & Prolog & $\surd$
\\\hline {\sc PowerAqua}~\cite{DBLP:conf/esws/LopezMU06} & graph representation
(query triples) & X\\\hline
{\sc DeepQA}~\cite{FerrucciBCFGKLMNPSW10} & semantic triples & $\surd$\\\hline
\end{tabular}
\caption{Semantic meaning representations}
\label{tab:semantic-meaning-representation}
\end{table*}

The parse tree (which may or not contain semantic information in parse tree
nodes) must then be {\it interpreted} in some internal semantic representation.
Table~\ref{tab:semantic-meaning-representation} displays semantic meaning
representations used in different systems.
As shown in this table, the internal semantic representation can be or not the
targetted query representation. 
The expressivity of the database query is discussed in section
section~\ref{sec:big-picture-expressivity}. 
The semantic representation is intented to capture as much as possible users'
intent, and is sufficient to generate the final database query. 
While the syntactic representation may contain lots of ambiguities, the semantic
representation does not. 
The semantic representation depends on the data structure. However, to
demonstrate their portability features, some systems contain components that
can translate the semantic representation for different database query
languages.

\subsubsection{Query languages}
Relational databases are generally associated with the SQL query language (and
temporal databases are associated to an extension of SQL called temporal SQL).
Multidimensional databases (e.g., data warehouses) are associated with MDX,
which is a SQL-like query languages dedicated for handling measures, dimensions
and hierarchies (key concepts of multidimensional models).
Other data stuctures are associated with their own query language. For instance,
hierarchical lists (or specification lists) is not associated with a language,
but to a similar structure, which is a template where empty slots correspond to
the expected items of the resultset. XML databases are associated to XPath, and
semantic databases (e.g., RDF) are usually queried with languages derived from
SPARQL which looks similar to SQL.
Recently, new query languages have appeared (``no-SQL''). 
\TODO{Develop\ldots}

\paragraph{Selection}
\label{sec:query-languages-selection}
a selection consists in choosing the expected database attribute to appear in
the resultset.
In SQL, the selection is expressed as \verb?SELECT t.x?, where $t$ is a table
and $x$ a field or attribute belonging to $t$. In SPARQL it corresponds to
\emph{variables} that are defined in the query, and that must satisfy the
constraints defined in the \verb?WHERE? section of the SPARQL query (see
section~\ref{sec:query-languages-constraints}).
In MDX, the selection corresponds to the ordered set of dimensions or measures
that should appear in the resultset, with the information about the level of the
expected attributes in the hierarchy, and the filters. For instance,
\verb?SELECT Country.[All members]? corresponds to the selection of the
dimension ``Country'' with no filter (i.e. selection of all members of the
dimension ``Country'').
Selection for a hierarchical list corresponds to a slot in the structure
template.
A selection in a XPath query corresponds to\ldots\footnote{\TODO{Find what}}
In SPARQL, a selection can be expressed also differently. Indeed, the output of
a SPARQL query can be a data structure of the same kind than the data themselves
(i.e. a graph). This is performed with the \verb?CONSTRUCT? keyword.

\paragraph{Constraints}
\label{sec:query-languages-constraints}
Canstraints in database queries aim at reducing the size of the resultset.
In the case of SQL, SPARQL and MDX, such constraints are introduced with the
\verb?WHERE? keyword. For SQL, these constraints can define how different tables
can be \emph{joined} together such that attributes belonging to different tables
can appear in a single view (i.e. in the resultset). Other kinds of constraints
are about the values of attributes.
For SPARQL, the constraints are expressed with \emph{triples} (which can be
made of some \emph{blank node} or undefined node). Thus, a SPARQL constraint is
about relations between entities (see \emph{joins} for relational databases), or
about entities themselves (which type must be the entity? If this is a
\emph{litteral}, in what range of value must it belong?,etc.).

\paragraph{Query modifiers}
\label{sec:query-languages-modifiers}
Conceptually, modifiers are primitives that modify the resultset afterwards.
Such primitives can - order the resultset along a given attribute (e.g.,
\verb?ORDER BY? in SQL); - select at most $n$ \emph{tuples} of the resultset
(e.g., \verb?LIMIT x? in SQL); - select only different tuples (e.g., the
keyword \verb?DISTINCT? in SQL).
In SPARQL, it is possible to test if a set of constraints can be satisfied from
the data, and to combine these tests with optional constraints (\verb?OPTIONAL?
keyword) or mandatory constraints (\verb?FILTER? keyword).

@@

\subsection{Domain-independence}
Domain-independance is the ability of a system, to operate accross different
domains simultaneously. 
In practice, this means that a domain-independant system would be on top of
different data sources that belong to different domains, and that this system is
able to meet users' requests for any of these domains. 
The main difficulty of building domain-independant systems is the ability to
translate NL constructions differently in the different application domains.
For instance, the qualitative expression ``middle-aged'' can be interpreted
completely differently accross the different domains.

@@

\subsubsection{Classic translation approach}

The classic translation approach consists in going from (1) to (4).
To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.
